Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-02-10 ribose-paper.RMD
moved figs

---
title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
author: "Kristian Lian, Daniel Hammarström, Stian Ellefsen, Håvard Hamarsland"
output:
  word_document:
    reference_docx: reference-template-02.docx
editor_options: 
  chunk_output_type: console
---

@@

Regular resistance exercise training (RT) increases muscle mass and muscle strength, due to the adaptations stimulated by  mechanical strain. Unfortunately, responses to RT are not uniform across the population. Studies have shown RT-induced muscle growth to vary widely between individuals, with as much as 10-15% showing considerably impaired growth (ALVAREZ18;MANN14;THALACKER13). This may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13). While tweaking training modalties and -protocols induces different responses in different populations (HAMMARSTROM20), little is known regarding optimisation of RT to individual phenotypes. Generally, meta-analyses favour a moderate volume of RT over low volume RT regarding muscle growth (KRIEGER09;RHEA03;SCHOENFELD17), substantiated by recent RCTs (HAMMARSTROM20;SCHOENFELD19). Although Hammarström and colleagues (2020) found that a moderate volume led to more prominent gains in muscle mass, approximately 50% of the participants did not experience true benefits of moderate volume compared to low, with a blunted muscle growth. Due to the variations observed in muscle growth responses despite standardising RT protocols (HAMMARSTROM20;SCHOENFELD17;STEC16), it is apparent that increasing RT volume alone does not convert low-responders to high-responders. Therefore, other means than RT *per se* seems necessary to circumvent this discrepancy.

Indeed, ingesting nutrients such as protein- and creatine-supplements effectively optimise RT (CERMAK12;LANHERS15;LANHERS17). However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT. This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis (MUL15;TANAKA18). Energy availability is important to regulate central processes of muscle growth, specifically ribosome biogenesis, a key factor for increased muscle protein synthesis (MPS) (FIGUEIREDO19;HAMMARSTROM20;VONWALDEN19;TANAKA18). Interestingly, variations in RT-induced responses have been linked to ribosome biogenesis, where participants responding poorly to RT also showed a blunted ribosome biogenesis (HAMMARSTROM20;STEC16). A blunted ribosome biogenesis negates responses to RT by failure to produce novel ribosomes, thus an inability to increase the skeletal muscle cell’s translational capacity (VONWALDEN19;FIGUEIREDO19). Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the regulation of ribosome biogenesis.

Ribosome biogenesis is suggested to be regulated by the mammalian target of rapamycin complex 1 (mTORC1) network (KUSNADI15;VONWALDEn19;TANAKA18). Stimulated by growth factors, hormones, mechanical loading and nutrients, mTORC1 governs energy distribution by turning on/off energy-consuming and energy-generating pathways, depending on cellular energy levels (HOPPE09). This greatly affects ribosome biogenesis, being a process that consumes approximately 80% of cellular energy and materials (VONWALDEN19;TANAKA18). Upon stimulation by mTORC1, assembly of the preinitiation complex (PIC) marks the initiation of ribosome biogenesis (VONWALDEN19;FIGUEIREDO19). The PIC consists of the RNA polymerase I complex, the SL-1 complex and the TIF-1A/RRN3 protein (FIGUEIREDO19, VONWALDEN19). Further, the upstream binding factor (UBF) and c-Myc have vital roles in the initiation, where UBF binds the PIC to the rDNA promoter and c-Myc as a transcription factor (FIGUEIREDO19, VONWALDEN19). Pol I transcribes the 47S pre-rRNA, suggested as the rate-limiting step in ribosome biogenesis (VONWALDEN19, FIGUEIREDO19). The 47S pre-rRNA is processed into the 18S, 5.8S and 28S rRNA (MULLINEUX12, VONWALDEN19), while the last rRNA 5S is transcribed by Pol III, and the ribosomal proteins by Pol II (MULLINEUX12, VONWALDEN19). Importantly, these events facilitates increased ribosomal content of each skeletal muscle cell, subsequently allowing a higher basal- and roof for MPS, suggested to predict long term adaptations to RT (HAMMARSTROM21). Since a blunted ribosome biogenesis has been observed to reduce/inhibit RT responses, despite increasing mechanical loading (HAMMARSTROM20), targeting hormonal and/or nutritional signals may ameliorate this. Previously, high glucose treatment promoted ribosome biogenesis compared to a repressed ribosome biogenesis with low glucose treatment and glucose starvation in mice and cell cultures (MARIAPPAN11;TANAKA15;ZHAI12), making glucose an interesting player in this perspective.

>>>>

---
title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
author: "Kristian Lian, Daniel Hammarström, Stian Ellefsen, Håvard Hamarsland"
bibliography: resources/ribose.bib
csl: resources/jap.csl
output:
  word_document:
    reference_docx: reference-template-02.docx
editor_options: 
  chunk_output_type: console
---

@@

Systematic resistance training (RT) increases muscle mass and strength due to repeated mechanical and metabolic stress. Unfortunately, responses to RT vary widely between individuals, with as many as 10-15% showing impaired growth (ALVAREZ18;MANN14;THALACKER13). Such inability to benefit from RT may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13).<!-- Check these refs, do they correspond to genet epigen and internal mileu? -->
Although RT program modifications such as exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020]<!-- Add ref for more rt program modifications -->
changes in e.g., training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020]. Therefore, other means than modification of RT variables seems necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory? -->

Nutritional supplements such as protein and creatine effectively optimize RT adaptations (CERMAK12;LANHERS15;LANHERS17). However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT. This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis (MUL15;TANAKA18). Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?  --> which in turn determines muscle growth by increasing the muscles translational capacity (FIGUEIREDO19;VONWALDEN19;TANAKA18;HAMMARSTROM20;STEC16). Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the regulation of ribosome biogenesis.

Ribosome biogenesis is suggested to be in part regulated by the mammalian target of rapamycin complex 1 (mTORC1) network (KUSNADI15;VONWALDEn19;TANAKA18). Stimulated by growth factors, hormones, mechanical loading and nutrients, mTORC1 governs energy distribution by turning on/off energy-consuming and energy-generating pathways, depending on cellular energy levels (HOPPE09). This greatly affects ribosome biogenesis, being a process that consumes approximately 80% of cellular energy and materials (VONWALDEN19;TANAKA18). Upon stimulation by mTORC1, assembly of the preinitiation complex (PIC) marks the initiation of ribosome biogenesis (VONWALDEN19;FIGUEIREDO19). The PIC consists of the RNA polymerase I complex, the SL-1 complex and the TIF-1A/RRN3 protein (FIGUEIREDO19, VONWALDEN19). Further, the upstream binding factor (UBF) and c-Myc have vital roles in the initiation, where UBF binds the PIC to the rDNA promoter and c-Myc as a transcription factor (FIGUEIREDO19, VONWALDEN19). Pol I transcribes the 47S pre-rRNA, suggested as the rate-limiting step in ribosome biogenesis (VONWALDEN19, FIGUEIREDO19). The 47S pre-rRNA is processed into the 18S, 5.8S and 28S rRNA (MULLINEUX12, VONWALDEN19), while the last rRNA 5S is transcribed by Pol III, and the ribosomal proteins by Pol II (MULLINEUX12, VONWALDEN19). Importantly, these events facilitates increased ribosomal content of each skeletal muscle cell, subsequently allowing a higher basal- and roof for MPS, suggested to predict long term adaptations to RT (HAMMARSTROM21). Since a blunted ribosome biogenesis has been observed to reduce/inhibit RT responses, despite increasing mechanical loading (HAMMARSTROM20), targeting hormonal and/or nutritional signals may ameliorate this. Previously, high glucose treatment promoted ribosome biogenesis compared to a repressed ribosome biogenesis with low glucose treatment and glucose starvation in mice and cell cultures (MARIAPPAN11;TANAKA15;ZHAI12), making glucose an interesting player in this perspective.
